---
title: "Something something"
author: "Josh Sumner"
subtitle: ''
always_allow_html: true
output:
  pdf_document:
    highlight: tango
    number_sections: no
    toc: yes
    toc_depth: '4'
  html_document:
    code_folding: hide
    highlight: textmate
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 4
    toc_float: no
  word_document:
    toc: yes
    toc_depth: '4'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "")
knitr::include_graphics

library(DT)
library(tidyverse)
library(RColorBrewer)
library(scales)
library(grDevices)
library(survival)
library(survminer)
library(KMsurv)
library(ggfortify)
library(lubridate)
library(broom)
library(cmprsk)
library(cowplot)
library(condsurv)

theme_set(theme_minimal() +
  theme(axis.line.y.left = element_line(),
        axis.line.x.bottom = element_line()))
#spColors<-c(brewer.pal(7, "Dark2"))
#swColors<-c(brewer.pal(12, "Paired"))

`%not_in%` <- purrr::negate(`%in%`)
```

```{r, eval=F}
#install packages that failed with install.packages("NAME") 
#if condsurv fails, use this block
library(remotes)
remotes::install_github("zabore/condsurv")
```


# Read In


```{r}
#df<-read_csv("strokes_4.csv")
df<-read_csv("stroke_smoke.csv") #smoking change
colnames(df)<-str_to_lower(str_replace_all(colnames(df), "\\s|-", "_"))
colnames(df)<-str_replace_all(colnames(df),"reference_event", "stroke_1")
colnames(df)<-str_remove_all(colnames(df), "_\\(smoking\\)")
df<-df%>%mutate(stroke_2 = ifelse(!is.na(stroke_2_condition_start_date_days_from_reference), 1,0))  #smoking change
colnames(df)
```

# Check Censoring

```{r}
df$count_censored<-apply(df,1,function(x) 
            sum(sapply("censored", function(codes) grepl("censored",x))))

list<-apply(df,2,function(x) 
            sum(sapply("censored", function(codes) grepl("censored",x))))
sort(list, decreasing=T)%>%
  head(5)

df%>%
  mutate(ifelse(count_censored>=1, 1, 0))%>%
  group_by(count_censored>=1)%>%
  summarize(n=n())

# df<-df%>%
#   mutate(across(
#     .cols = everything(), #maybe change this, something like starts_with("malignant"),
#     .fns = ~ str_replace(.,"censored", "missing")
#   ))
```


# Cleaning

## Make Survival Data

```{r}
isNeg<-function(x){ifelse(x < 0, 1, x)}

cleaned<-df%>%
  rowid_to_column("id")%>%
  mutate(primary_race = ifelse(grepl("black", primary_race, ignore.case = T), "Black", primary_race))%>%
  filter(primary_race %in% c("Black", "White"))%>%
  rowwise()%>%
  mutate(TIME = ifelse(is.na(stroke_2_condition_start_date_days_from_reference) & 
                         is.na(death_date_of_death_days_from_reference) &
                         is.na(c_embolism_condition_start_date_days_from_reference), 
                      #7049, #what time should be recorded as event time if no event happened? last time we see them.
                       round(max(c(hypertension_admission_start_date_days_from_reference,
                           ace_inhib_order_start_date_days_from_reference,
                           diuretics_order_start_date_days_from_reference,
                           beta_blockers_order_start_date_days_from_reference,
                           cc_blocker_order_start_date_days_from_reference,
                           arb_order_start_date_days_from_reference,
                           chd_condition_start_date_days_from_reference), na.rm=T), 4),
                       round(min(c(stroke_2_condition_start_date_days_from_reference, death_date_of_death_days_from_reference, c_embolism_condition_start_date_days_from_reference), na.rm=T), 4)))%>%
  mutate(TIME = ifelse(TIME == Inf, 7049, TIME))%>%
  mutate(TIME = isNeg(TIME))%>%
  mutate(status = 
    ifelse(!is.na(death_date_of_death_days_from_reference) &
          TIME==round(death_date_of_death_days_from_reference, 4) , 2, #death
    ifelse(!is.na(stroke_2_condition_start_date_days_from_reference) &
          TIME==round(stroke_2_condition_start_date_days_from_reference, 4) , 1, #stroke
    ifelse(!is.na(c_embolism_condition_start_date_days_from_reference) &
          TIME==round(c_embolism_condition_start_date_days_from_reference, 4) , 1, #stroke
    0))))%>% #censoring
  mutate(TIME = ifelse(TIME==0, 0.5, TIME))%>%
  rename(ref_age = stroke_1_age_at_event,
         bmi = bmi_average_calculated_bmi,
         stroke_2_time = stroke_2_condition_start_date_days_from_reference,
         death_time = death_date_of_death_days_from_reference,
         embolism_time = c_embolism_condition_start_date_days_from_reference)%>%
  mutate(across(
    .cols = contains("days_from_reference"),
    .fns = ~ isNeg(.),
    .names = "{col}_time"))%>%
  mutate(across(
    .cols = contains("_time"),
    .fns = ~ ifelse(is.na(.), 7049, .)))%>%
  mutate(bmi = ifelse(is.na(bmi), median(bmi, na.rm=T), bmi))%>%
  mutate(BMI = ifelse(is.na(bmi), "censored", #from "missing"
                      ifelse(bmi < 18.5, "Underweight",
                      ifelse(18.5 <= bmi & bmi < 25, "Normal",
                             ifelse(25 <= bmi & bmi < 30, "Overweight", "Obese")))))%>%
  mutate(BMI = as.factor(BMI))%>%
  select(id, race=primary_race, ref_age, bmi, TIME, status, contains("_time"), BMI, stroke_2, smoke_tobacco_status)%>% #from stroke_2_condition
  ungroup()
  
colnames(cleaned)<-str_remove_all(colnames(cleaned), "_start_date_days_from_reference_time")

cleaned<-cleaned%>%
  mutate(BB = ifelse(beta_blockers_order < TIME, 1, 0),
         CC = ifelse(cc_blocker_order < TIME, 1, 0),
         DIU = ifelse(diuretics_order < TIME, 1, 0),
         ACE = ifelse(ace_inhib_order < TIME, 1, 0),
         HYP = ifelse(hypertension_admission < TIME, 1, 0),
         ARB = ifelse(arb_order < TIME, 1, 0),
         CHD = ifelse(chd_condition < TIME, 1, 0))%>%
  mutate(smoke = ifelse(is.na(smoke_tobacco_status), "Unknown",
                                ifelse(smoke_tobacco_status == "Unknown If Ever Smoked", "Unknown",
                                ifelse(smoke_tobacco_status == "Never Assessed", "Unknown",
                                ifelse(smoke_tobacco_status == "Never Smoker", "Nonsmoker",
                                ifelse(smoke_tobacco_status == "Passive Smoke Exposure - Never Smoker", "Nonsmoker",
                                       "Smoker"))))))%>%
  select(id, ref_age, TIME, status, bmi, BMI, hypertension_admission, HYP, ace_inhib_order, ACE, diuretics_order, DIU, beta_blockers_order, BB, cc_blocker_order, CC, arb_order, ARB, chd_time = chd_condition, CHD, race, smoke)
head(cleaned)
```



### Make Time Dependent Survival Data

```{r}
switchBinary<-function(x){-1*(x-1)}
td<-tmerge(
    data1 = cleaned %>% select(id, TIME, status, ref_age, BMI, bmi, race, smoke), 
    data2 = cleaned %>% select(id, TIME, status, BB, beta_blockers_order), 
    id = id, 
    stroke2 = event(TIME, status),
    beta = tdc(BB)
    )
#nrow(td)
td<-tmerge(
  data1 = td,
  data2 = cleaned%>%select(id, TIME, status, CC, cc_blocker_order),
  id = id,
  cc = tdc(CC)
)
#nrow(td)
td<-tmerge(
  data1 = td,
  data2 = cleaned%>%select(id, TIME, status, HYP, hypertension_admission),
  id = id,
  hyp = tdc(HYP)
)
#nrow(td)
td<-tmerge(
  data1 = td,
  data2 = cleaned%>%select(id, TIME, status, ACE, ace_inhib_order),
  id = id,
  ace = tdc(ACE)
)
#nrow(td)
td<-tmerge(
  data1 = td,
  data2 = cleaned%>%select(id, TIME, status, ARB, arb_order),
  id = id,
  arb = tdc(ARB)
)
#nrow(td)
td<-tmerge(
  data1 = td,
  data2 = cleaned%>%select(id, TIME, status, CHD, chd_time),
  id = id,
  chd = tdc(CHD)
)
td<-td%>%
  mutate(across(
    .cols = c(beta, cc, hyp, ace, arb, chd),
    .fns = ~switchBinary(.)
  ))%>%
  mutate(across(
    .cols = c(BMI, race, smoke, beta, cc, hyp, ace, arb, chd),
    .fns = ~as.factor(.)
  ))
```

# Data Dictionary

## Cleaned

`cleaned` Data:

`id`: Row identifier for unique simulated subjects

`ref_age`: Age at first stroke

`TIME`: Time to second stroke in days. When a time was not present this defaults to 7049.

`status`: 
status = 0 : censored
status = 1 : stroke
status = 2 : Death by competing risk

`bmi`: Numeric BMI

`BMI`: Discretized BMI
Underweight: bmi under 18.5
Normal: bmi between 18.5 and 25
Overweight: bmi between 25 and 30
Obese: bmi over 30
missing: bmi was not found

`hypertension_admission`: Time of hypertension diagnosis (moved to 1 if diagnosis occured before first stroke)

`HYP`: Binary yes/no (1/0) whether patient was diagnosed with hypertension.

`ace_inhib_order`: Time of ace inhibitor order (moved to 1 if order occured before first stroke)

`ACE`: Binary yes/no (1/0) whether patient has ACE inhibitors at any point before their second stroke.

`diuretics_order`: Time of diuretics order (moved to 1 if order occured before first stroke)

`DIU`: Binary yes/no (1/0) whether patient is on diuretics at any point before their second stroke.

`beta_blockers_order`: Time of beta blockers order (moved to 1 if order occured before first stroke)

`BB`: Binary yes/no (1/0) whether patient is on beta blockers at any point before their second stroke.

`cc_blocker_order`: Time of calcium channel blockers order (moved to 1 if order occured before first stroke)

`CC`: Binary yes/no (1/0) whether patient is on calcium channel blockers at any point before their second stroke.

`arb_order`: Time of agents acting on renin-angiotensin system order (moved to 1 if order occured before first stroke)

`ARB`: Binary yes/no (1/0) whether patient is taking agents acting on the renin-angiotensin system at any point before their second stroke.

`chd_time`: Time of CHD diagnosis (moved to 1 if order occured before first stroke)

`CHD`: Binary yes/no (1/0) whether patient has been diagnosed with CHD at any point before their second stroke.

`smoke`: Smoking status reduced to 3 levels, unknown, smoker, and nonsmoker.

`race`: Subject race.

## TD

`td` Data:

`id`: Row identifier for unique simulated subjects, subjects may have multiple rows of data if their time dependent variables changed during their time 'on the study'.

`ref_age`: Age at first stroke

`TIME`: Time to second stroke in days. When a time was not present this defaults to 7049.

`status`: 
status = 0 : censored
status = 1 : stroke
status = 2 : Death by competing risk

`bmi`: Numeric BMI

`BMI`: Discretized BMI
Underweight: bmi under 18.5
Normal: bmi between 18.5 and 25
Overweight: bmi between 25 and 30
Obese: bmi over 30
missing: bmi was not found

`race`: Subject race.

`smoke`: Smoking status reduced to 3 levels, unknown, smoker, and nonsmoker.

`tstart`: Start of timeframe for patient with given set of time dependent covariates

`tstop`: Stop of timeframe for the patient with given set of time dependent covariates

`stroke2`: Binary whether event occured during this timeframe for this patient

`beta`: Binary whether or not whether subject had beta blockers in this timeframe

`cc`: Binary whether or not whether subject had calcium channel blockers in this timeframe

`hyp`: Binary whether or not whether subject had diagnosed hypertension in this timeframe

`ace`: Binary whether or not whether subject had ace inhibitors in this timeframe

`arb`: Binary whether or not whether subject had agents acting on renin-angiotensin system in this timeframe

`chd`: Binary whether or not whether subject had CHD in this timeframe


# Models

```{r}
td%>%
  group_by(hyp, stroke2)%>%
  summarize(n=n())
```


```{r}
tdnd<-td%>%
  mutate(age_group = factor(ifelse(ref_age < 45, "<45",
                            ifelse(ref_age<60, "<60",
                            ifelse(ref_age <75, "<75", 
                                   ">75")))))%>%
  mutate(stroke2=ifelse(stroke2 %in% c(1,2), 1,0)) #stroke or death = 1, censor = 0

tdcxph<-coxph(Surv(time = tstart, time2 = tstop, event = stroke2) ~ age_group+smoke+hyp+chd, data = tdnd)

tdcxph %>% broom::tidy(exp=T)
s<-summary(tdcxph)
conc<-s$concordance[1]
assumption<-cox.zph(tdcxph)
assumption
aic1<-extractAIC(tdcxph)
aic1
plotMod1<-ggsurvplot(survfit(tdcxph),
           data= tdnd,
           censor = F,
           censor.shape= 124,
           censor.size = 1.75,
           title = "Model 1 (Classic Covariates)",
           size = 0.75,
           legend="none")
plotMod1$plot<-plotMod1$plot+
  ggplot2::annotate("label", x=2000, y=0.25, label=paste0("Concordance: ", round(conc, 4), "\n AIC:", round(aic1[2], 2)))
plotMod1
ggsave("mod_1.png", width = 6, height=4)
```



```{r}
tdcxph2<-coxph(Surv(time = tstart, time2 = tstop, event = stroke2) ~ beta+cc+arb+ace+BMI+strata(race), data = tdnd)

tdcxph2 %>% broom::tidy(exp=T)
s<-summary(tdcxph2)
conc2<-s$concordance[1]
cox.zph(tdcxph2)
aic2<-extractAIC(tdcxph2)
aic2
plotMod2<-ggsurvplot(survfit(tdcxph2),
           data= tdnd,
           censor = F,
           censor.shape= 124,
           censor.size = 1.75,
           title = "Model 2 (New Covariates)",
           size = 0.75,
           legend="top")
plotMod2$plot<-plotMod2$plot+
  ggplot2::annotate("label", x=2000, y=0.25, label=paste0("Concordance: ", round(conc2, 4), "\n AIC:", round(aic2[2],2)))
plotMod2
ggsave("mod_2.png", width = 6, height=4)
```

```{r}
tdcxph3<-coxph(Surv(time = tstart, time2 = tstop, event = stroke2) ~ age_group+strata(race)+BMI+beta+cc+ace+arb+smoke+chd, data = tdnd)
# age_group + race + BMI + beta + cc + arb + ace + smoke + chd

#tdcxph3 %>% broom::tidy(exp=T)
s<-summary(tdcxph3)
conc3<-s$concordance[1]
#cox.zph(tdcxph3)
aic3<-extractAIC(tdcxph3)
aic3
plotMod3<-ggsurvplot(survfit(tdcxph3),
           data= tdnd,
           censor = F,
           censor.shape= 124,
           censor.size = 1.75,
           title = "Model 3 (Combination)",
           size = 0.75,
           legend="top")
plotMod3$plot<-plotMod3$plot+
  ggplot2::annotate("label", x=2000, y=0.25, label=paste0("Concordance: ", round(conc3, 4), "\n AIC:", round(aic3[2],2)))
plotMod3
ggsave("mod_3.png", width = 6, height=4)
```

```{r}
anova(tdcxph, tdcxph2, tdcxph3)
```



```{r}
fit <- list(Original = survfit(tdcxph), New = survfit(tdcxph2), Mixture = survfit(tdcxph3))
```


```{r, fig.height=6}
graph<-ggsurvplot(fit=fit,
           data = tdnd,
           combine = TRUE,
           censor = FALSE,
           title = "Survival Curves of All Models",
           tables.theme = theme_cleantable(),  # Clean risk table
           palette = "jco")

ggsave("combined_models_tab.png", height=2, width=7)
```


```{r, fig.height=8, fig.width = 9}
ggsurvplot(data = tdnd, 
           fit = survfit(tdcxph3),
           xlab = "Years",
           xscale = 365.25,
           break.x.by = 730.5,
           size = 0.75,
           censor = TRUE,
           censor.shape= 124,
           censor.size = 1.75,
           fun = "cumhaz",
           legend.title = "",
           legend = "bottom", 
           risk.table = T,
           risk.table.y.text = T)
```



# Time Dependent AUC

Run this at your own risk, it seems to be RAM heavy

```{r, eval=F}
install.packages('survAUC')
```

```{r, eval=F, echo=F}
library(survAUC)
```

```{r, eval=F, echo=F}
rio::export(cleaned, "cleaned_strokes.csv")
rio::export(tdnd, "time_dependent.csv")
```

```{r, eval=F, echo=F}
cleanedcsv<-read_csv("cleaned_strokes.csv")
set.seed(987)
cleaned_train<-cleanedcsv%>%
  slice_sample(prop = 0.75)

cleaned_test<-anti_join(cleanedcsv, cleaned_train)




cxphT1<-coxph(Surv(time = TIME, event = status) ~ ref_age+smoke+HYP+CHD+race+bmi, data = cleaned_train)

# tdcxphT3<-coxph(Surv(time = tstart, time2 = tstop, event = stroke2) ~ age_group+strata(race)+BMI+beta+cc+ace+arb+smoke+chd, data = tdnd_train)

pred1<-predict(cxphT1)
pred2<-predict(cxphT1, newdata=cleaned_test)

Surv.rsp<-Surv(cleaned_train$TIME, cleaned_train$status)
Surv.rsp.new<-Surv(cleaned_test$TIME, cleaned_test$status)

times <- seq(0, 8000, 100)

AUC_CD <- AUC.cd(Surv.rsp, Surv.rsp.new, pred1, pred2, times)

AUC_CD
```





```{r}
ggsurvplot(
    fit = survfit(Surv(TIME, status) ~ strata(race), data = cleaned), 
    size = 0.75,
    censor = FALSE, #turns off the vertical tickmarks at censoring times
    legend.labs = c("Black", "White"),
    xlab = "Days", 
    ylab = "Overall survival probability")+#layering works per normal gg language
  ggtitle("Kaplan Meier Curve")
```













## Competing Risk Regression



```{r}
ci_fit<-cuminc(ftime = cleaned$TIME,
               fstatus = cleaned$status, #event 1 = stroke , event 2 = CR (death)
               group = cleaned$race,
               cencode = 0)
ci_fit
ci_fit[["Tests"]]

ggcompetingrisks(
  fit = ci_fit, 
  multiple_panels = FALSE,
  xlab = "Days",
  ylab = "Cumulative incidence of Events",
  title = "Stroke vs Death",
  ylim = c(0, 1)
)
```

```{r}
ciplotdat <- 
  ci_fit %>% 
  list_modify("Tests" = NULL) %>% 
  map_df(`[`, c("time", "est"), .id = "id")%>% 
  mutate(id = recode(
    id, 
    "Black 1" = "Black:Stroke", 
    "Black 2" = "Black:Death", 
    "White 1" = "White:Stroke",
    "White 2" = "White:Death")
    ) %>% 
  separate(id, c("Group", "Event"), ":") 


ggplot(ciplotdat, aes(x = time, y = est, color = Group)) +
  geom_step(aes(linetype = Event), size = 0.5)  +
  ylim(c(0, 1))+
  labs(x = "Days", 
       y = "Cumulative incidence",
       title = "Stroke and Death by Race") +
  annotate("text", x = 0, y = 1, hjust = 0,
           label = paste0(
             "Stroke p = ", 
             ifelse(ci_fit$Tests[1, 2] < .001, 
                    "<.001", 
                    round(ci_fit$Tests[1, 2], 3)))) + 
  annotate("text", x = 0, y = 0.92, hjust = 0,
           label = paste0(
             "Competing Risk (Death) p = ", 
             ifelse(ci_fit$Tests[2, 2] < .001, 
                    "<.001", 
                    round(ci_fit$Tests[2, 2], 3)))) +
  theme_classic() +
  theme(plot.title = element_text(size = 14),
        legend.title = element_blank(),
        legend.position = "bottom") 
ggsave("competing_risks.png", width=6, height=4)
```



```{r}
csh_1_fit <- 
  coxph(
    Surv(TIME, ifelse(status == 1, 1, 0)) ~ race + ref_age, 
    data = cleaned
    )

broom::tidy(csh_1_fit, exp = TRUE)

cz <- cox.zph(csh_1_fit)
print(cz)
plot(cz)
```

## Citations

```{r}
citation("dplyr")
citation("survival")
citation("cmprsk")
citation("survAUC")
```


```{r}
sessionInfo()
```











